---
title: "petrify"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gtfsrealtime)
library(gtfstools)
library(lubridate)
library(purrr)
library(stringr)
library(data.table)
library(tidyverse)
library(ggunc)
library(petrifygtfsrealtime)
```

## Read data

### Read static GTFS

```{r}
static = read_gtfs(Sys.getenv("GTFS_PATH"))
```

### Read realtime data

Data are saved with filenames in UTC, figure out what files we need for a day in local time.

```{r}
start = ymd_hms("2026-06-02 03:00:00", tz = "America/Los_Angeles")

start_utc = with_tz(start, "Etc/UTC")
end_utc = start_utc + period(1, "day")

rt_files = map(seq(start_utc, end_utc, by="day"), function (date) {
  path = file.path(Sys.getenv("GTFS_RT_PATH"), format(date, "%Y/%m/%d"), "vehicle-positions")
  stopifnot(dir.exists(path))
  candidates = list.files(path, full.names = TRUE)

  # get the date-time from the filename
  dates = lubridate::ymd_hms(str_extract(candidates, "[0-9]{14}"))

  return(candidates[dates >= start_utc & dates <= end_utc])
}) |> unlist()

realtime = map(rt_files, \(x) read_gtfsrt_positions(x, "America/Los_Angeles")) |>
  rbindlist()
```

## Remove duplicates

The same position update may appear in multiple files. Fix.

```{r}
realtime = realtime[,.SD[1], by=.(vehicle_id, timestamp)]
```

## Petrify GTFS

This is the core function of the package.

```{r}
p = petrify(realtime, static)
```

## Stringline plots

A [stringline plot](https://pvibien.com/stringline.htm) is a very useful tool for debugging the output. Here, we make a plot for the LA Metro 720 to make sure it looks as expected. These plots are dense, so it is often useful to save as a PDF for ability to pan and zoom.

```{r}
patterns = gtfstools::get_stop_times_patterns(p)
stringline(p, patterns, "720-13196", directionid = 0)
ggsave("720.pdf", width=8, height=14)
```

## Write GTFS for further analysis

```{r}
write_gtfs(p, "la_metro_test.zip")
```